### 2018-11-08 paper/complexity.tex
New proof for the upper bound of test executions

The argument so far has shown that reference process $\pmax$ requires the most test 
executions among all CSP processes $P$ whose collections of minimal hitting sets  
$\minhits(P/s)$ are never empty for any trace $s$, so that the branch 
(\ref{eq:ufb}) of the related test cases $U_F(j)$ can never be entered.  
Lemma~\ref{lemma:failureshittingsets} implies that an empty collection 
$\minhits(P/s)$
is equivalent to $(s,A)\in\fails(P)$ for all $A\subseteq\Sigma$. This is equivalent to
$\maxrefs(P/s) = \{ \Sigma \}$ or $\minaccs(P/s) = \{ \varnothing\}$.

It remains to 
consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
may require more test executions for their associated tests $U_F(j)$ than the bound
given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider an {\it arbitrary} process  $Z$ represented by its transition graph $G(Z)$ with $p$ nodes. 
Let $v$ be any node of $G(Z)$, such 
that $\minaccs(v)\neq\{\varnothing \}$. Now modify $G(Z)$ by setting $v$'s acceptance label
to $\{\varnothing \}$. This produces a modified node $\ol v$ with $[\ol v]^0=[v]^0$, but
$\minaccs(\ol v)=\{\varnothing \}$. Now it is checked whether $\ol v$ is equivalent to any other node of $G(Z)$. If this is not the case, the modified transition graph $G(\ol Z)$ constructed by
exchanging $v$ with $\ol v$ and keeping all other nodes and all edges of $G(Z)$  in the new graph without changes has exactly the same number of nodes as $G(Z)$.

Now consider a test case ${\ol U}_F(j)$ for $j\in\{ 0,\dots,pq-1 \}$ derived from $G(\ol Z)$
and compare it to $U_F(j)$ derived from $G(Z)$. Let $s$ be a trace of  ${\ol U}_F(j)$ such that
$\#s\le j$ and   $G(\ol Z)/s = \ol v$. Then we have two cases to distinguish.
\begin{enumerate}
\item If $\#s < j$, then ${\ol U}_F(j)/s$ allows for one additional execution step entering branch 
(\ref{eq:ufb}) of the test, due to construction of $\ol v$.

\item If $\#s = j$, then ${\ol U}_F(j)/s$ loses $\card{\minhits(v)}$ execution steps checking 
that $Z/s$ never blocks one of the hitting sets in $\minhits(v)$ and wins one execution step 
because again, branch 
(\ref{eq:ufb}) of the test can now be entered. The difference $1 - \card{\minhits(v)}$ has 0 as
the largest possible value. This applies, for example, if the intersection of 
all minimal acceptances of $n$ contains exactly one element.
\end{enumerate}

As a very rough estimate for the maximal number of additional test executions of ${\ol U}_F(j)/s$
in comparison to $U_F(j)$ we can therefore assume that $\card{\minhits(v)} = 1$ and all
$\card{\Sigma}^k$ traces of length $k < j$ lead to the new node $\ol v$. As a result,
${\ol U}_F(j)/s$ has one additional execution step for every trace of length less than $j$.  
This results in an upper bound of 
\begin{equation}
\sum_{i=0}^{j-1}n^i = \frac{n^j - 1}{n - 1} \quad\text{with}\ n = \card{\Sigma}
\end{equation}
test executions. Summing up over additional test executions stemming from al test cases 
 ${\ol U}_F(j),\ j = 0,\dots,pq-1$ results in the following upper bound for the number of 
 additional executions.
\begin{equation}
\sum_{j=0}^{pq-1}\frac{n^j - 1}{n - 1} = \frac{n^{pq} - npq + pq - 1}{(n-1)^2}\quad\text{with}\ 
n=\card{\Sigma}.
\end{equation}
This expression has the order of magnitude $O(n^{pq-2})$.  Therefore, the additional test executions won by being able to enter branches (\ref{eq:ufb}) in test executions  can be neglected in comparison to the order of magnitude calculated for $\pmax$ in (\ref{eq:maxO}).

These considerations lead to the following theorem
%
\begin{theorem}
Given a process alphabet $\Sigma$ with cardinality $
\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
the normalised transition graph of $P$ has $p$ states, and the fault domain 
${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then the maximal 
number of test executions to be performed   using the complete test 
suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} is of order
%
\begin{equation*} 
O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
\end{equation*} 
\xbox
\end{theorem}
%
In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
non-empty subset of $\Sigma$ whose events cannot be completely refused in a
given process state of the reference model; this leads to a worst-case
estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
offered to the SUT in the last step of the test execution. This number is significantly larger than the worst-case estimate $\binom{n}{\lfloor n/2\rfloor}$ calculated above for the 
hitting sets to be checked.
In Fig.~\ref{fig:minhita}, the reduction is visualised by plots of
the two functions.

>>>>

The argument so far has shown that reference process $\pmax$ requires the most test 
executions among all CSP processes $P$ whose collections of minimal hitting sets  
$\minhits(P/s)$ are never empty for any trace $s$, so that the branch 
(\ref{eq:ufb}) of the related test cases $U_F(j)$ can never be entered.  
Lemma~\ref{lemma:failureshittingsets} implies that an empty collection 
$\minhits(P/s)$
is equivalent to $(s,A)\in\fails(P)$ for all $A\subseteq\Sigma$. This is equivalent to
$(s,\Sigma)\in\fails(P)$ or
$\maxrefs(P/s) = \{ \Sigma \}$ or $\minaccs(P/s) = \{ \varnothing\}$.

It remains to 
consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
may require more test executions for their associated tests $U_F(j)$ than the bound
given for $\pmax$ in (\ref{eq:pmaxcomplexity}), because process states $Z/s$ with  
$\minhits(Z/s)=\varnothing$ allow for test executions entering branch (\ref{eq:ufb}).
To this end, consider a test cases $U_F(j)$ constructed from such a process $Z$. Every trace
$s\in\trc(Z)$ with $\#s<j$ ending in a process state $Z/s$ with $\minhits(Z/s)=\varnothing$ allows for 
%
\begin{itemize}
\item one execution of branch (\ref{eq:ufb}), where the test execution 
$(Z\parallel[\Sigma]U_F(j))/s$ stops after $\epass$, and
\item $\card{[Z/s]^0}$   continuations of the test execution with events $e\in [Z/s]^0$. 
\end{itemize}
%
For every trace $s\in\trc(Z)$ with $\#s=j$,
%
\begin{itemize}
\item one execution of branch (\ref{eq:ufb}) follows if 
$\minhits(Z/s)=\varnothing$, and otherwise
\item $\card{\minhits(Z/s)}$ executions checking acceptance of minimal hitting sets. 
\end{itemize}
%
For a rough estimate of the worst case upper bound suppose that
%
\begin{enumerate}
\item $[Z/s]^0 = \Sigma$ for all traces $s$ of $Z$,
\item all traces $s$ with $\#s<j$ end in a state with empty minimal hitting sets, and
\item all traces $s$ with $\#s=j$ end in a state with a maximal number $\binom{n}{n/2}$
 of hitting sets.
\end{enumerate}
%
Note that this situation cannot be realised for all $j\in\{0,\dots,pq-1\}$, because
the traces of 
$U_F(p-1)$ already cover all states of $Z$'s transition graph according to Lemma~\ref{lemma:extendV}, and if all states of $Z$ have empty hitting sets, there are no acceptance checks to be performed in the last step of the test execution. Therefore, the   upper bound calculated next cannot be reached by a real CSP process.

With the 3 assumptions above, we calculate that
%
\begin{itemize}
\item  $U_F(0)$ has  $\binom{n}{n/2}$ executions,
\item $U_F(j),\ j > 0$ has $\sum_{i=0}^{j-1} n^i = \frac{n^j - 1}{n-1}$ executions of branch (\ref{eq:ufb}), ($n = \card{\Sigma}$), and  
\item $U_F(j),\ j > 0$ has $\binom{n}{n/2}\cdot n^j$ executions where the acceptance of 
hitting sets is checked after having run through a trace of length $j$. 
\end{itemize}
%
Summing up over all $U_F(j)$ for $j=0,\dots,pq-1$ this results in an upper bound of
%
\begin{equation}
\frac{(n-1)n(n^{pq} - 1)\binom{n}{\lfloor\frac{n}{2}\rfloor} + n^{pq} - n^2(pq-1) + n(pq-2)}{(n-1)^2n}
\end{equation}
%
test executions. Since this upper bound cannot be reached anyway, 
we just calculate its order of magnitude, and this results again in $O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}  \cdot n^{pq-1}\big)$, as calculated already for $\pmax$ in (\ref{eq:maxO}).

Summarising these complexity calculations, this results in the following theorem.
%
\begin{theorem}
\label{th:maxexecs}
Given a process alphabet $\Sigma$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
the normalised transition graph of $P$ has $p$ states, and the fault domain 
${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most 
$q\ge p$ states. Then the maximal 
number of test executions to be performed   using the complete test 
suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} is of order
%
\begin{equation*} 
O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
\end{equation*} 
For processes $P$ satisfying $(s,\Sigma)\not\in\fails(P)$ for all traces $s$, the reachable
precise  upper bound is given by
%
\begin{equation*}\label{eq:pmaxcomplexity} 
\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}.
\end{equation*}
\xbox
\end{theorem}
%

%
In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
non-empty subset of $\Sigma$ whose events cannot be completely refused in a
given process state of the reference model; this leads to a worst-case
estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
offered to the SUT in the last step of the test execution. This number is significantly larger than the worst-case estimate $\binom{n}{\lfloor n/2\rfloor}$ calculated above for the 
hitting sets to be checked.
In Fig.~\ref{fig:minhita}, the reduction is visualised by plots of
the two functions.
